---
title: "Reproducible Reports Assignment - Course Project 1"
output: html_document
---
The data in this assignment, provided by course instructors, was from a personal activity monitoring device that collects data at 5-minute intervals. The data was downloaded from the course website. The variables were: 1) steps, number of steps taken at 5-min intervals; 2) date, date measurement was taken, and 3) interval,  indentifier for the 5-min intervals.  R codes to read and analyze the data, along with analysis results are presented below. 

```{r}
# read data
setwd("C:/Users/user/DataScience_Coursera/repdata-data-activity/")
activity <- read.csv("activity.csv", stringsAsFactors=FALSE, header=TRUE)

#format variables
activity$date <- as.Date(as.character(activity$date))

# exclude missing values
activity_complete <- na.exclude(activity)
```
# What is the mean total number of steps taken per day?
```{r}
total_steps_per_day <- aggregate(activity_complete$steps, by=list(date=activity_complete$date), sum)

colnames(total_steps_per_day) <- c("date", "total.steps")
```
# 1. Total number of steps per day
```{r}
total_steps_per_day
```

# 2. Histogram of total number of steps per day
```{r}
hist(total_steps_per_day$total.steps, col="blue", xlab="Number of steps per day", main="Histogram of\n total number of steps taken each day")
```

# 3.Mean and median total number of steps per day
```{r}
summary(total_steps_per_day$total.steps)
```
# What is the average daily activity pattern?
```{r}
average_steps_per_interval <- aggregate(activity_complete$steps, by=list(interval=activity_complete$interval), mean)

colnames(average_steps_per_interval) <- c("interval", "average.steps")

plot(x=average_steps_per_interval$interval, y=average_steps_per_interval$average.steps, type='l', col=2, xlim=c(0,max(average_steps_per_interval$interval)), 
     xlab="Time interval (min)", ylab="Average number of steps",
     main="Time series plot average number of steps")

```

# The 5-minute interval that displays the max average number of steps: 
```{r}
average_steps_per_interval[which.max(average_steps_per_interval$average.steps),]
```

# Imputing missing values
# 1. Number of missing values
```{r}
sum(is.na(activity))
```

#2. creating a new data set with mean imputation
```{r}
require(Hmisc)
activity$imputed.steps <- with(activity, impute(steps, mean))
```

# 3. Histogram of total number of steps per day \n in imputed data set
```{r}
total_number_steps_a_day <- aggregate(activity$imputed.steps, by=list(date=activity$date), sum)

colnames(total_number_steps_a_day) <- c("date", "total.steps")

hist(total_number_steps_a_day$total.steps, col="blue", xlab="Number of steps per day", main="Histogram of total number of steps \n taken each day (imputed data set)")
```

# 3.Mean and median total number of steps per day in imputed data set
```{r}
summary(total_number_steps_a_day$total.steps)

```

Yes, imputation produced different values compared to the data set earlier. The imputed data set contained additional data points between 5,000 & 10,000 steps.  As seen in the histogram, the imputed data set has additional data points between the first quartile and the median. 

## Activity patterns between weekdays and weekends
```{r}
weekdays.list <- c("Monday", "Tuesday", "Wednesday", "Thursday","Friday")

activity$date.group <- c("weekends", "weekdays")[(weekdays(activity$date) %in% weekdays.list)+1L]

activity$date.group <- as.factor(activity$date.group)

ggplot(activity, aes(interval, steps)) + 
        geom_line(stat="summary", fun.y="mean") + 
        facet_grid(. ~ date.group) +
        ylab("steps") +
        xlab("interval") +
        ggtitle("Average number of steps in imputed data set")
```

yes, activity patterns between weekdays and weekends are different. In general, weekday acttivities data are skewed to the left where as weekend activities are somewhat normally distributed.